\documentclass[../../solutions]{subfiles}
\title{Template}
\author{Julian Gilbey}
\begin{document}
\maketitle

% \settheorem{1}{2}{3}
% \begin{lemma}
%   
% \end{lemma}
% \popthm

\setexercise{4}{2}{1}
\begin{exercise}
  Prove that any pair of adjoint functors $\afunc{F}{G}{\sC}{\sD}$
  restrict to define an equivalence between the full subcategories
  spanned by those objects $c\in\sC$ and $d\in\sD$ for which the
  components of the unit $\eta_c$ and of the counit $\epsilon_d$,
  respectively, are isomorphisms.
\end{exercise}

\begin{proof}
  Let $\sC'$ and $\sD'$ be the respective full subcategories.

  We must first show that $Fc\in\sD'$ whenever $c\in\sC'$ and
  $Gd\in\sC'$ whenever $d\in\sD'$.  Suppose $c\in\sC'$, so
  $\func{\eta_c}{c}{GFc}$ is an isomorphism.  We must show that
  $\func{\epsilon_{Fc}}{FGFc}{Fc}$ is an isomorphism.  We have
  $\epsilon_{Fc}\cdot F\eta_c=1_{Fc}$ by the triangle identities, and
  $F\eta_c$ is an isomorphism (as functors preserve isomorphisms), so
  $\epsilon_{Fc}=(F\eta_c)^{-1}$ is also an isomorphism.  Thus
  $Fc\in\sD'$ as required.

  Similarly, assume that $d\in\sD'$ so that
  $\func{\epsilon_d}{FGd}{d}$ is an isomorphism; we must show that
  $\eta_{Gd}$ is an isomorphism.  We have
  $G\epsilon_d\cdot\eta_{Gd}=1_{Gd}$ by the triangle identities, so
  $\eta_{Gd}=(G\epsilon_d)^{-1}$ is an isomorphism.

  Thus $F$ and $G$ restrict to functors $\sC'\rightleftarrows\sD'$; we
  also denote these restricted functors by $F$ and~$G$.

  Finally, we must show that $\afunc{F}{G}{\sC'}{\sD'}$ gives an
  equivalence of categories.  We have the natural transformations
  $\Func{\eta}{1_\sC}{GF}$ and $\Func{\epsilon}{FG}{1_\sD}$.  Since
  every component of $\eta$ and~$\epsilon$ is an isomorphism, $\eta$
  and~$\epsilon$ are natural isomorphisms, so the restriction of $F$
  and~$G$ to these full subcategories defines an equivalence of
  categories.
\end{proof}

\end{document}

