\documentclass[../../solutions]{subfiles}
\title{Template}
\author{}
\begin{document}
\maketitle

% \settheorem{1}{2}{3}
% \begin{lemma}
%   
% \end{lemma}
% \popthm

\setexercise{4}{1}{1}
\begin{exercise}
  Show that functors $\func{F}{\sC}{\sD}$ and $\func{G}{\sD}{\sC}$ and
  bijections $\sD(Fc, d) \cong \sC(c, Gd)$ for each $c \in \sC$ and
  $d \in \sD$ define an adjunction if and only if these bijections
  induce a bijection between commutative squares (4.1.4). That is,
  prove Lemma 4.1.3.
\end{exercise}

\begin{proof}
  Let $\sC$ and $\sD$ be categories and let $\func{F}{\sC}{\sD}$,
  $\func{G}{\sD}{\sC}$ be functors.

  Suppose that we are given, for any $c\in \sC$, $d \in \sD$, a
  bijection
  $$\varphi_{c,d}\colon \sD(Fc, d) \cong \sC(c, Gd)
  $$
  Denote its inverse by $\psi_{c,d}$.  We will omit the subscripts if
  they are understood.

  Recall that this collection of bijections forms an adjunction (i.e.,
  is natural in $c$ and $d$) if and only if the following identities
  hold:
  \begin{equation}
    \label{eq:4-1-1-adj1}
    \varphi(k\cdot f^\sharp) = Gk \cdot f^\flat
  \end{equation}
  for any $c \in \sC$, $d, d' \in \sD$, $\func{k}{d}{d'}$ and
  morphisms as pictured below
  $$
  \begin{tikzcd}
    Fc
    \ar[r, "f^\sharp"]
    \ar[dr]
    & d
    \ar[d, "k"]
    \\
    & d'
  \end{tikzcd}
  \hspace{1cm}
  \begin{tikzcd}
    c
    \ar[r, "f^\flat"]
    \ar[dr]
    & Gd
    \ar[d, "Gk"]
    \\
    & Gd'
  \end{tikzcd}
  $$
  where $(f^\sharp, f^\flat)$ is an adjoint pair; and similarly
  \begin{equation}
    \label{eq:4-1-1-adj2}
    \psi(g^\flat \cdot h) = g^\sharp \cdot Fh
  \end{equation}
  for any $c,c'\in \sC$, $d \in \sD$, $\func{h}{c}{c'}$ and morphisms
  as pictured below
  $$
  \begin{tikzcd}
    Fc
    \ar[d, "Fh"']
    \ar[dr]
    \\
    Fc'
    \ar[r, "g^\sharp"']
    & d'
  \end{tikzcd}
  \hspace{1cm}
  \begin{tikzcd}
    c
    \ar[d, "h"']
    \ar[dr]
    \\
    c'
    \ar[r, "g^\flat"']
    & Gd'
  \end{tikzcd}
  $$
  where $(g^\sharp, g^\flat)$ is an adjoint pair.

  Our goal, as per Lemma 4.1.3, is to show that the equations in
  \eqref{eq:4-1-1-adj1} and \eqref{eq:4-1-1-adj2} hold if and only if
  the left square below commutes precisely when the right square
  commutes:
  $$
  \begin{tikzcd}
    Fc
    \ar[r, "f^\sharp"]
    \ar[d, "Fh"']
    & d
    \ar[d, "k"]
    \\
    Fc'
    \ar[r, "g^\sharp"']
    & d'
  \end{tikzcd}
  \hspace{1cm}
    \begin{tikzcd}
    c
    \ar[r, "f^\flat"]
    \ar[d, "h"']
    & Gd
    \ar[d, "Gk"]
    \\
    c'
    \ar[r, "g^\flat"']
    & Gd'
  \end{tikzcd}
  $$
  Indeed, suppose first that the equations in \eqref{eq:4-1-1-adj1}
  and \eqref{eq:4-1-1-adj2} hold.  Suppose that the left square
  commutes. Then we have
  $$
  Gk \cdot f^\flat = \varphi(k \cdot f^\sharp) = \varphi(g^\sharp Fh)
  = g^\flat h
  $$
  where we use commutativity of the left square and the identities in
  \eqref{eq:4-1-1-adj1}, \eqref{eq:4-1-1-adj2}.  So the right square
  commutes.

  Similarly, if the right square commutes, then we have
  $$
  k \cdot f^\sharp = \psi(Gk \cdot f^\flat) = \psi (g^\flat h) =
  g^\sharp \cdot Fh
  $$
  so the left square commutes.

  Conversely, suppose that the left square commutes exactly when the
  right square commutes.  Then in particular, in the context of
  \eqref{eq:4-1-1-adj1} we can choose $c' = c$, $h = 1_c$,
  $g^\flat =Gk \cdot f^\flat$, making the right square commute by
  construction.  Then the left square commutes as well.  But
  $F(1_c)=1_{Fc}$ and
  $g^\sharp = \psi(g^\flat) = \psi(Gk\cdot f^\flat)$, so commutativity
  of the left square means $\psi(Gk\cdot f^\flat) = k\cdot f^\sharp$.
  Since $\varphi$ is an inverse of $\psi$, this yields equation
  \eqref{eq:4-1-1-adj1}.

  Similarly, in the context of \eqref{eq:4-1-1-adj2} we can choose
  $d = d'$, $k = 1_d$ and $f^\sharp = g^\sharp \cdot Fh$.  Then the
  left square commutes by construction, hence the right square
  commutes, which precisely amounts to equation \eqref{eq:4-1-1-adj2}.
\end{proof}

\end{document}

